WSClean: an implementation of a fast, generic wide-field imager

### 2014-02-05 articles/2014/Article-WSClean/wsclean.tex
Improved WSSC benchmark and channel benchmark, work on article

@@

\section{Application of WSClean to the MWA}
The MWA correlator splits observations in data products of typically a few minutes
TODO

TODO: calculate when to use w-snapshots

TODO: mention full-polarization imaging

\section{Discussion} \label{sec:discussion}
WSClean has been build out of necessacity to be able to make images with the MWA.

@@

\section{Optimal gridding kernel size} \label{sec:gridding-appendix}
\begin{figure}
\begin{center}
\includegraphics[width=8cm]{img/benchmark-kernelsize/kernel}
\caption{...}
\label{fig:timing-kernelsize}
\end{center}
\end{figure}
As discussed in \S\ref{sec:gridding}, WSClean uses a Kaiser-Bessel window against aliasing. The type of window function has no effect on the gridding performance, but the size of the kernel affects gridding performance quadratically. Fig.~\ref{fig:timing-kernelsize} shows that this quadratic relation becomes significant for kernel sizes $\gtrapprox 10$. Decreasing the kernel size to values below $7$ has little effect on performance, because for small kernels the visibility-reading rate is lower than the gridding rate. We have not noticed much benefit of larger kernels except in rare cases where a bright source lies just outside the imaged field of view. Therefore, WSClean uses a default of 7 pixels for the gridding kernel. It can be increased when necessary. Different hardware might have a slightly different optimum because of different reading and calculation performance.

>>>>

TODO: derive function for $t_\textrm{CASA} = t_\textrm{WSClean}$.

@@

\section{Application of WSClean to the MWA}
For technical reasons, the MWA correlator splits observations in data products of typically a few minutes. This simplifies $w$-snapshot imaging, which can be done by phase rotating each snapshot to zenith with a possibly additional phase-centre shift.

TODO: calculate when to use w-snapshots
TODO: calculate optimum splitting time?

\subsection{Beam correction}
Because the MWA consists of fixed, beamformed dipole antennas, the voltage beam of a MWA tile is described by a complex, non-diagonal Jones matrix. Alternatively, Mueller matrices can be used as well, but in general these result in more computations \citep{revisiting-me-i}. When assuming all individual antennas of the tiles are working properly, all tiles have the same beam. The beam can then be corrected in image space for all tiles at once with
\begin{equation}\label{eq:beam-correction}
 I(l, m) = J(l,m)^{-1} I'(l, m) \left(J(l,m)^*\right)^{-1},
\end{equation}
where each term is a $2\times2$ complex matrices, with $J$ the voltage beam and $^*$ denoting the conjugate transpose. Because $J$ is complex and non-diagonal, all 4 combinations of the cross-correlated polarizations $p$ and $q$\footnote{The instrumental polarizations of the MWA are often referred to as $X$ and $Y$ because many software treats the two polarizations as such. However, this is not correct because $X$ is officially defined as being parallel to the celestial equator.} including the imaginary part of the $pq$ and $qp$ are required to calculate any of the Stokes parameters. The $pp$ and $qq$ have no imaginary part due to the $uv$-symmetry. To make proper MWA beam correction possible, WSClean can output real valued $pp$, $pq$, $qp$, $qq$ images and imaginary $pq$ and $qp$ images. Each polarization requires a separate run of the $w$-stacking algorithm. Once these six images are created, a separate program is used to perform the correction of Eq.~\eqref{eq:beam-correction}. This method avoids expensive beam-correcting kernels during gridding, but only works because all tiles have the same beam. To apply the same method on heterogeneous arrays such as LOFAR, each set of correlations with a different combination of station beams will have to be imaged separately, which will increase the cost of the algorithm. The AWImager uses an intermediate method, and corrects a common dipole factor in image space and the phased-array beam factor in $uv$-space, which allows the gridding kernel to be smaller compared to correcting both in $uv$-space \citep{awimager-2013}.

\section{Discussion} \label{sec:discussion}
WSClean has been build out of necessacity to be able to make images with the MWA, because no

@@

\section{Optimal gridding kernel size} \label{sec:gridding-appendix}
\begin{figure}
\begin{center}
\includegraphics[width=8cm]{img/benchmark-kernelsize/kernel}
\caption{Effect of antialiasing kernel size on performance.}
\label{fig:timing-kernelsize}
\end{center}
\end{figure}
As discussed in \S\ref{sec:gridding}, WSClean uses a Kaiser-Bessel window against aliasing. The type of window function has no effect on the gridding performance, but the size of the kernel affects gridding performance quadratically. Fig.~\ref{fig:timing-kernelsize} shows that this quadratic relation becomes significant for kernel sizes $\gtrapprox 10$. Decreasing the kernel size to values below $7$ has little effect on performance, because for small kernels the visibility-reading rate is lower than the gridding rate. We have not noticed much benefit of larger kernels except in rare cases where a bright source lies just outside the imaged field of view. Therefore, WSClean uses a default of 7 pixels for the gridding kernel. It can be increased when necessary. Of course, different hardware might give slightly different results because of different reading and calculation performance.
